---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

if(!require(librarian)) install.packages("librarian")
pkgs <- c("targets",
          "sf",
          "leaflet",
          "sf",
          "ggplot2",
          "patchwork",
          "dplyr")
librarian::shelf(pkgs)
```

# warmwaterfish

<!-- badges: start -->
<!-- badges: end -->

The goal of `warmwaterfish` is to look at the prevalence of warm-water fish species in coastal Nova Scotia under climate change.

## Data Pipeline

The data pipeline is built using the `targets` package. We use `targets` to avoid re-querying GBIF every time we want to run the script (i.e. it will only re-query if we change the query it). For a quick tutorial on targets, see Jaimie Harbin's [demo](https://github.com/AtlanticR/bioRworkshops/blob/master/22_targets/targets.Rmd) or the targets [website](https://books.ropensci.org/targets/).

The main points to know are that the code for the pipeline is in `_targets.R` and the data are stored in the `_targets/objects` folder. To run the entire pipeline, just run `targets::tar_make()` in the R console. To see the status of the pipeline, run `targets::tar_visnetwork()`. To load the output of a target into your R session, run `targets::tar_load(<target_name>)` or `targets::tar_load_everything()`.

```{r}
tar_load(c(inatdata,maritimes))
tar_visnetwork()
```

## Example Plots

Species counts:
```{r}
ggplot(inatdata)+
  geom_bar(aes(species))+
  #rotate x-labels 90
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))+
  theme_minimal()
```

User contributions:
```{r}
ggplot(inatdata)+
  geom_bar(aes(recordedBy))+
  #rotate x-labels 90
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))+
  theme_minimal()
```

Spatial distribution of observations:
```{r}
p1 <- ggplot(inatdata)+
  geom_bar(aes(round(decimalLatitude*4)/4))+
  coord_flip()+
  theme_minimal()

p2 <- ggplot(maritimes)+
  geom_sf()+
  geom_sf(data= inatdata, aes(color=species), size=2, alpha=0.7)+
  theme_minimal()
p1 + p2
```

Heatmap of observation density:
```{r, warning=FALSE}

ggplot(maritimes) +
  geom_sf() +
  stat_sf_coordinates(data = inatdata) +
  geom_hex(data = inatdata, 
           aes(x = decimalLongitude, y = decimalLatitude), 
           alpha = 0.7, 
           bins = 30) +
  scale_fill_viridis_c(name = "Sightings\nDensity") +
  facet_wrap(~species) +
  theme_minimal()
```

